Cache is King: Presenting a Suite of Fast, Correlation Function Codes

### 2015-01-23 paper/ms.tex
Added an users-guide

\lstset{
  language=C,                % choose the language of the code
  numbers=left,                   % where to put the line-numbers
  stepnumber=1,                   % the step between two line-numbers.
  numbersep=10pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  tabsize=2,                      % sets default tabsize to 2 spaces
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=true,         % sets if automatic breaks should only happen at whitespace
}

@@

\lstinputlisting[firstline=11]{../xi_theory/include/cellarray.h}

@@

\subsection{Hand-written Vectorization Support}
Advanced Vector Extensions (AVX) has been available in cpu's more recent than 2011. AVX allows the processing of 8 floats or 4 double simultaneously -- thus, 
potentially increasing the throughput by a factor of 8x/4x. However, automatic vectorization is not always possible by the compiler and in those cases 
we can write AVX vector intrinsics to directly manipulate 8 floats/4 doubles.

\section{Benchmarks \& Scaling}
In this section we present the runtimes and OpenMP scaling for the different codes.

\section{Conclusions}

>>>>

\lstset{
  language=C,                % choose the language of the code
  basicstyle=\ttfamily,      % sets font
  numbers=left,                   % where to put the line-numbers
  numberstyle=\small, 
  stepnumber=1,                   % the step between two line-numbers.
  numbersep=10pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  tabsize=2,                      % sets default tabsize to 2 spaces
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=true,         % sets if automatic breaks should only happen at whitespace
}

@@

\begin{lstlisting}
typedef struct{
  DOUBLE *x;
  DOUBLE *y;
  DOUBLE *z;
  int64_t nelements;
} cellarray;
\end{lstlisting}

@@

\subsection{Hand-written Vectorization Support}
Advanced Vector Extensions (AVX) has been available in cpu's more recent than 2011. AVX allows the processing of 8 floats or 4 double simultaneously -- thus, 
potentially increasing the throughput by a factor of 8x/4x. However, automatic vectorization is not always possible by the compiler and in those cases 
we can write AVX vector intrinsics to directly manipulate 8 floats/4 doubles\footnote{Another option would be to use the vectorclass written by Agner Fog here: \url{http://agner.org/optimize/vectorclass/}}.

\section{Benchmarks \& Scaling}
In this section we present the runtimes and scalings for different number of particles, \rmax and OpenMP threads for the codes. 
For all of the scaling tests, only an auto-correlation calculation was used and the fiducial catalog contains $\sim 1.2$ million 
galaxies on a periodic cube of side 420 \hMpc.

\subsection{Scaling with Number of Particles}
In Fig.~\ref{fig:scaling_numpart}, we show the scaling for the three codes with the number of particles. For this scaling, we subsampled 
the fiducial mock to attain 10 logarithmic steps in particle number ranging from $1.2\times10^4$ to $1.2\times10^6$.

\begin{figure}[htbp]
\includegraphics[clip=true,width=\linewidth]{timings_Mr19_numpart}%
\caption{Scaling with particle number for \xir, \xirppi, \wprp }
\label{fig:scaling_numpart}
\end{figure}

\subsection{Scaling with \rmax}
\begin{figure}[htbp]
\includegraphics[clip=true,width=\linewidth]{timings_Mr19_rmax}%
\caption{Scaling with \rmax for \xir, \xirppi, \wprp }
\label{fig:scaling_rmax}
\end{figure}

\subsection{Scaling with OpenMP threads}
\begin{figure}[htbp]
\includegraphics[clip=true,width=\linewidth]{timings_Mr19_openmp}%
\caption{OpenMP scaling for \xir, \xirppi, \wprp }
\label{fig:scaling_openmp}
\end{figure}

\section{Conclusions}
I have presented a suite of three fast correlation function codes that take advantage of the underlying hardware. The reasons why the codes 
are much faster for typical workloads are:
\begin{itemize}
\item A `bin-lattice' scheme is used to first partition the computational domain into 3-D cells that can then be used to prune majority of the volume. 
\item To obtain better cache-locality, the particle list is duplicated into a contiguous array for each dimension. Thus, all particles that fall into the
same 3-D cell are stored in a contiguous array. 
\item With the AVX instruction set, modern CPU's can process 8 floats/4 doubles simultaneously. The codes contain hand-written AVX intrinsics that offer a 
factor of few speedup compared to the compiler generated scalar code. 
\end{itemize}
